---
title: "Lab 09 - More inference"
subtitle: "Due: Friday, Apr 17 at 11:59pm EST"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: "labs.css"
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

# Introduction

The data are a sample of 500 observations in a study where air pollution at a 
road is related to traffic volume and meteorological variables. All data was
collected by the Norwegian Public Roads Administration. 

# Getting started

Clone your assignment repo into RStudio Cloud and open the R Markdown file.
Don't forget to load in the necessary packages and configure git.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(infer)
library(broom)
```

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="your name", user.email="your email")
```

If you would like your git password cached for a week for this project, type 
the following in the Terminal:

`git config --global credential.helper 'cache --timeout 604800'`

You will need to enter your GitHub username and password one more time after 
caching the password.This is only good for this single RStudio Cloud project; 
you will need to cache your credentials for each project you create.

# Data

The response variable (column 1) consist of hourly values of the logarithm 
of the concentration of $\mbox{NO}_2$ (particles), measured at 
Alnabru in Oslo, Norway. 
The predictor variables (columns 2 to 8) 
are the logarithm of the number of cars per hour, temperature 2 meter above 
ground (degree C), wind speed (meters/second), the temperature difference 
between 25 and 2 meters above ground (degree C), wind direction 
(degrees between 0 and 360), hour of day and day number from October 1, 2001.

```{marginfigure}
StatLib---Datasets Archive. (2020). Lib.stat.cmu.edu. Retrieved 7 April 2020, 
from http://lib.stat.cmu.edu/datasets/
````

```{r read_data, eval=FALSE}
no <- read_delim(file = "http://lib.stat.cmu.edu/datasets/NO2.dat",
                 delim = "\t", col_names = FALSE)
```

# Set a seed

Before you get started, set the seed so each person in your group will
be able to reproduce your analysis.

```{r}
set.seed(8675309)
```

# Exercises

In addition to the linear model fitting functions, statistical distribution 
functions, and base R arithmetic functions, your code should only contain 
functions from the loaded R packages above unless explicitly stated in an 
Exercise.

## Data cleaning and exploration

1. Fix the column names in `no` to something meaningful.

2. Create a well-polished visualization that investigates the relationship 
   between log concentration of $\mbox{NO}_2$ and the hour of the day. 

3. Comment on what you observe from your visualization in Exercise 2.
   
## CLT-based inference

4. Create a 95% confidence interval for the mean log concentration of 
   $\mbox{NO}_2$.
   What assumptions must hold for this interval estimate to be valid?

5. Let $\mu$ be the mean number of cars per hour that travel on Alnabru in Oslo, 
   Norway. Suppose the average number of cars per hour
   that travel on similar roads is known to be 2000. Is the sample data in
   `no` strong enough evidence at the $\alpha = .05$ level to suggest that 
   the road in Alnabru is less traveled?
   
6. Assume all population parameters are unknown. In a general hypothesis 
   testing framework, what is the smallest observed
   test statistic value, $\frac{\bar{x} - \mu_0}{s / \sqrt{n}}$, we could 
   obtain and still reject the null hypothesis
   at the $\alpha = 0.01$ significance level when $H_A: \mu > \mu_0$ and $n=32$.
   Here quantity $\mu_0$ represents the value of $\mu$ under the null 
   hypothesis.
   
7. If you reject the null hypothesis at the $\alpha = 0.02$ significance level, 
   then you will also reject the null hypothesis at the $\alpha = 0.01$
   significance level. Explain in detail, or with an example, why this
   claim is true or false.

## Regression inference

8. Consider a linear model with log concentration of $\mbox{NO}_2$ as the 
   response and temperature 2 meter above ground (degree C) as the single 
   predictor. Compare the simulation-based inference 90% confidence interval 
   for the slope and the 90% confidence interval for the slope derived from the
   regression output. You may assume the linear regression conditions hold 
   without verifying via plot diagnostics.

9. Perform backwards elimination with AIC as the selection criteria, where the
   full model consists of predictors in columns 2 through 7 of `no` and the
   response is log concentration of $\mbox{NO}_2$. Display the model terms, 
   estimates    and p-values in a tidy format. *Hint*: sort the data 
   chronologically before model fitting, it will make things easier in 
   Exercise 10.
   
10. Check if all the linear model conditions are satisfied for the final model
    from Exercise 9.
   
11. Compute 95% confidence intervals for two coefficients in your final
    model from Exercise 9. Provide an interpretation for these intervals.

# Submission

Knit to PDF to create a PDF document. Stage and commit all remaining changes,
and push your work to GitHub. Make sure all files are updated on your 
GitHub repo.

Please only upload your PDF document to Gradescope. Associate the “Overall” 
graded section with the first page of your PDF, and mark where each answer 
is to the exercises. If any answer spans multiple pages, then mark all pages.

Only one team member needs to submit for the group. After you hit submit, 
go to View or edit group and select all your team members from the drop-down
menu.

# References

- StatLib---Datasets Archive. (2020). Lib.stat.cmu.edu. Retrieved 7 April 2020, 
  from http://lib.stat.cmu.edu/datasets/