---
title: "Lab 08 - Simulation-based inference"
subtitle: "Due: Friday, Apr 10 at 11:59pm EST"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: "labs.css"
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

# Introduction

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in the state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We'll work with a random sample of 
observations from this data set.

# Getting started

Clone your assignment repo into RStudio Cloud and open the R Markdown file.
Don't forget to load in the necessary packages and configure git.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(infer)
library(openintro)
```

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="your name", user.email="your email")
```

If you would like your git password cached for a week for this project, type 
the following in the Terminal:

`git config --global credential.helper 'cache --timeout 604800'`

You will need to enter your GitHub username and password one more time after 
caching the password.This is only good for this single RStudio Cloud project; 
you will need to cache your credentials for each project you create.

# Data

We'll read in `ncbirths` with

```{r read_nc_births, eval=FALSE}
data(ncbirths)
```

In the dataset we have observations on 13 different variables, some categorical 
and some numerical. The meaning of each variable is as follows.

variable         | description
---------------- | ---------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

# Set a seed

Before you get started, set the seed so each person in your group will
be able to reproduce your analysis.

```{r}
set.seed(71189752)
```

# Exercises

In addition to `quantile()`, `diff()`, and `is.na()`, your code should only 
contain functions from the loaded R packages above unless explicitly stated 
in an Exercise.

## Baby weights

```{marginfigure}
Wen, Shi Wu, Michael S. Kramer, and Robert H. Usher. "Comparison of birth weight distributions between Chinese and Caucasian infants." American Journal of Epidemiology 141.12 (1995): 1177-1187.
````

A 1995 study suggests that the average weight of Caucasian babies born in the 
U.S. is 3,369 grams (7.43 pounds). In this dataset we only have information on 
mother's race, so we will make the simplifying assumption that babies of 
Caucasian mothers are also Caucasian. 

We want to evaluate whether the average weight of Caucasian babies in NC has 
changed from the 1995 study by performing a simulation-based hypothesis test 
at the .05 significance level.

1. Write out the hypotheses for this test in words.

2. Write out the hypotheses for this test in symbolic notation.

3. Create a well-labelled plot of the null distribution and shade the
   p-value.

```{r}
sample_mean = ncbirths %>% 
   filter(whitemom == "white") %>% 
   summarise(avg_weight = mean(weight)) %>% 
   pull()
```

```{r}
null_dist = ncbirths %>% 
   filter(whitemom == "white") %>% 
   specify(response = weight) %>% 
   hypothesize(null = "point", mu=7.43) %>% 
   generate(reps = 5000, type = "bootstrap") %>% 
   calculate(stat = "mean")
```

```{r}
null_dist %>% 
   visualise() +
   shade_p_value(obs_stat = sample_mean,
                 direction = "two_sided")
```


4. Compute the p-value and interpret your findings within the context of the
   data and research question.

```{r}
null_dist %>% 
   get_p_value(obs_stat = sample_mean,
               direction = "two_sided")
```
Since p-value is less than the confidence level, 0.05, we reject the null hypothesis.

 
5. Compute a 95% confidence interval for the population mean weight of 
   Caucasian babies born in NC. Does this interval cover the value of $\mu$
   under your specified null hypothesis in Exercise 2? Explain why it does
   or does not.

```{r}
ncbirths %>% 
   filter(whitemom == "white") %>% 
   specify(response = weight) %>% 
   generate(reps = 5000, type = "bootstrap") %>% 
   calculate(stat = "mean") %>% 
   summarise(lb = quantile(stat, 0.025),
             ub = quantile(stat, 0.975))
```


The confidence interval does not include the mu, 7.43. The confidence interval should be consistent with the hypothesis testing, and they should be equivalent.


## Baby weight vs. smoking

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

6.  Make side-by-side boxplots displaying the relationship between `habit` and 
    `weight`. What does the plot highlight about the relationship between these 
    two variables?


```{r}
ncbirths %>% 
   ggplot(aes(x=habit,y=weight)) +
   geom_boxplot()
```
The median weight of the babies seems to be lower from moms who smoke than those from moms who do not smoke.


7. Before moving forward, save a version of the dataset omitting observations 
   where there are NAs for `habit`. You can call this version 
   `ncbirths_habitgiven`.

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following to 
first group the data by the `habit` variable, and then calculate the mean
`weight` in these groups using.


```{r}
ncbirths_habitgiven = ncbirths %>% 
   filter(!is.na(habit))
```

```{r eval=FALSE}
ncbirths_habitgiven %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

There is an observed difference based on our sample data, but we want to know 
if the difference is statistically significant. In order to answer this 
question we will conduct a hypothesis test.



8. Write the hypotheses for testing if the average weights of babies born to 
   smoking and non-smoking mothers are different. Define the difference as
   smoking - non-smoking.

H0: mu-smoking = mu-nonsmoking
Ha: mu=smoking != mu-nonsmoking



9. Perform an appropriate hypothesis test at the 0.10 significance level, 
   calculate the p-value, and interpret the results in the context of the data 
   and research question.

```{r}
mean_diff = ncbirths_habitgiven %>% 
   group_by(habit) %>% 
   summarise(mean_weight = mean(weight)) %>% 
   pull(mean_weight) %>% 
   diff()
```


```{r}
ncbirths_habitgiven %>%
   specify(response = weight, explanatory = habit) %>%
   hypothesize(null = "independence") %>%
   generate(1000, type = "permute") %>%
   calculate(stat = "diff in means", 
             order = c("smoker", "nonsmoker")) %>% 
   get_p_value(obs_stat = mean_diff, direction = "two_sided")
   
```

   
10. Given your conclusion in Exercise 9, which type of error could you possibly
   have made. What does this mean in the context of the testing problem.
   
Type 1 error: since we rejected the null hypothesis. but the null hypothesis could be true.   


11. In this testing framework, what is the probability of a type I error?

0.1 = alpha


# Submission

Knit to PDF to create a PDF document. Stage and commit all remaining changes,
and push your work to GitHub. Make sure all files are updated on your 
GitHub repo.

Please only upload your PDF document to Gradescope. Associate the “Overall” 
graded section with the first page of your PDF, and mark where each answer 
is to the exercises. If any answer spans multiple pages, then mark all pages.

Only one team member needs to submit for the group. After you hit submit, 
go to View or edit group and select all your team members from the drop-down
menu.

# References

- Wen, Shi Wu, Michael S. Kramer, and Robert H. Usher. "Comparison of birth 
  weight distributions between Chinese and Caucasian infants." American 
  Journal of Epidemiology 141.12 (1995): 1177-1187.