---
title: "Lab 03 - Tidy Graphics and Teamwork"
subtitle: "Due: Monday, Feb 3 at 11:59pm"
author: "Bin Han"
date: "1/31/2020"
output:
  pdf_document: default
  html_document: default
---

## Packages

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
```

## Exercise 1

## Exercise 2

```{r}
nerrs <- read_csv("data/nerrs.csv")
```

## Exercise 3

```{r}
ggplot(data=nerrs, aes(x=compound, y=concentration))+
    geom_bar(stat = "identity")+
    facet_grid(~date)+
    coord_flip()
```


```{r}
ggplot(data=nerrs, aes(x=compound, y=concentration, fill=date))+
    geom_bar(stat = "identity")+
    coord_flip()
```

```{r}
ggplot(data=nerrs, aes(x=reorder(compound, concentration), 
                       y=concentration, 
                       fill=date)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=concentration), 
            position=position_dodge(width=0.9),
            hjust=-0.1,
            size=3) +
  coord_flip()
```


## Exercise 4

1. The orientation of the graph makes it hard to compare between dates. It would 
be better to orientate the two dates horizontally so that for each compound, we 
can compare horizontally instead of vertically.

2. The color is based on the compounds, which might be overwhelming. It would be
better if we put compound on the axis and color them based on date.

## Exercise 5



*****

## Final checks

1. Did every team member commit?
2. Do all code chunks have names?
3. Are your figures appropriately sized?
4. Is all code and narrative following the 80 characters rule?
5. Did you proofread your narrative and run spell checker (button next to save)? 

*****



